---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# omicser

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

NDCN -Omics Browser (omicser TBC)
CZI, NDCN -Omics Browser
Data exploration playground for general -omics sharing.  Developed as part of the NDCN Open Science initiative Data Science Pilot March-Sept 202.


## Installation

You can install the current version of omicser from giuthub with:

`devtools::install_github("omicser")`

```{r example}
x <- 1

```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example2}
library(omicser)
## basic example code
```
 `devtools::build_readme()` 

```{r}



```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/master/examples>.

