---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# NDCN -Omics Browser (name TBC)

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->


## Introduction

The goal of this package is to allow exploration of -omics data using
an interactive "playground" (e.g., Shiny application),
accessible through an interactive web browser.

> *Note*:  although the NDCN browser is not yet officially named, there are references to the browser/package as `omicser` or `omxr`.

There are two types of users for this package:

- **Curators**: data analysts who configure the app and prepare datasets for viewing. These individuals should be fairly proficient with R coding, including familiarity with building Shiny apps and knowledgeable about -omics data structures.
- **Browsers**: scientists who are interested in exploring -omics data using the application. They do not require extensive knowledge about R, but may need to have it installed if they are running the app on their own computer.

The diagram below summarizes the steps required to use this package:

![]image_FIXME.jpg

Support for the following _post-QC_ -omics data types has been tested:
- _transcript_-omics
- _prote_-omics
- _metabol_-omics
- _lipid_-omics

The package is able to provide the following outputs:
- images visualizing data
- subsets of data
- web application that can be shared with collaborators

## Installing this package

```{r, 00-install, eval=FALSE}
devtools::install_github("ergonyc/omicser")
```

## Installation and usage for Browsers

Once a curator has prepared the data and configured the application,
the app can be launched from R:

```{r, 01-launch, eval=FALSE}
    library(omicser)
    omicser::launchApp()
```

For more information,
please see [FIXME]()
or consult with the Curator for your application.

## Installation and usage for Curators

A set of example scripts and a quickstart guide are available in the [`quickstart/`](quickstart/) directory of the repo.  The [Quickstart Guide](quickstart/README.md) will outline the flow from installation through data exploration.  i.e.

1. **Set up environment**: underlying tools/packages from R and python. [link](quickstart/01_environment_setup.md)
2. **Install software**: creates the browser and curation functions as an R package from github [link](quickstart/02_install.md)
3. **Curate data**: the _-omics_ data curated into a _database_ [link](quickstart/03_data_curation.md)
4. **Configure application**: connecting the _-omics database_ to the browser app [link](quickstart/04_configuration.md)
5. **Browse data**: explore the data!  [link](quickstart/05_browsing.md)
6. **Share application**: just a couple extra curation steps to share the data with others [link](quickstart/06_sharing.Rmd)

## Background and development

The NDCN Browser (`Omicser`) was developing using [`{golem}`](https://github.com/ThinkR-open/golem): "an opinionated framework for building production-grade shiny applications." 

Primary development was initiated as part of the NDCN Open Science Initiative Data Science Pilot March-Sept 2021, with hopes of formenting ongoing open-source development.

## Roadmap

Please view our [issues](https://github.com/ergonyc/omicser/issues),
especially those tagged as [high priority](https://github.com/ergonyc/omicser/issues?q=is%3Aopen+is%3Aissue+label%3A%22high+priority%22),
for more information about immediate plans for development.

+ data table visualization
+ QC report tab
+ Consistent `roxygen2` headers.  Currently importing functions (e.g. `require(“package”)`) or using the `package::function` syntax is inconsistent.  
+ Migrate all dplyr::  tidyverse table manipulation to data.frame syntax and/or dtdplyr.  
(test speed improvements on mac, with caveat of apple disabled multithreading)


```{r, render-quickstart1-6, eval=TRUE, include = FALSE}
rmarkdown::render("quickstart/01_environment_setup.Rmd")
rmarkdown::render("quickstart/02_install.Rmd")
rmarkdown::render("quickstart/03_data_curation.Rmd")
rmarkdown::render("quickstart/04_configuration.Rmd")
rmarkdown::render("quickstart/05_browsing.Rmd")
rmarkdown::render("quickstart/06_sharing.Rmd")
```


