---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# NDCN -Omics Browser (name TBC)

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->


## Introduction

This package allows creation of a browser app for exploring a variety of _post-QC_ -omics data.  Support for _transcript_-omics, _prote_-omics, _metabol_-omics and _lipid_-omics has been tested.  

The goal is to create a "playground" for exploring -omics data, with an aim towards sharing. Primary development was initiated as part of the NDCN Open Science initiative Data Science Pilot March-Sept 2021, with hopes of formenting ongoing open-source developmnet.

There are two types of "users" mode for this package.  First, a "curator" scientist who must prepare their dataset for browsing and perform some configuration so the app can access teh data.   Second, in "browser" mode the curated database(s) can be explored with the intention of developing intuition about the data and developing hypothesis to pose for the next steps of analysis.  

*Note:  although the NDCN browser is not yet officially named, there are references to the browser/package as `omicser` or `omxr`.  

## Installing this package

```{r, plug_mode, eval=FALSE}
devtools::install_github("ergonyc/omicser")
```

## Usage

Once installed two more steps are nescessary:
1. Data Curation <quickstart/03_data_curation.Rmd>
2. Browser Configuration  <quickstart/04_configurationn.Rmd>

After configuration the app is launched from R as:

```{r, plug_mode, eval=FALSE}
    library(omicser)
    omicser::launchApp()
```

## Quickstart

A set of examples scripts and a quickstart guide are available in the `quickstart/` directory of the repo.  The Quickstart Guide will outline the flow from installation through data exploration.  i.e.

1. Environment Setup: underlying tools/packages from R and pyhon <quickstart/01_environment_setup.Rmd>
2. Install: creates the browser and curation functions as an R package from github <quickstart/02_install.Rmd>
3. Data Curation: the _-omics_ data curated into a _database_  <quickstart/03_data_curation.Rmd>
4. Configuration: connecting the _-omics database_ to the browser app <quickstart/04_configurationn.Rmd>
5. Browse: explore the data!  <quickstart/05_browse.Rmd>


## Background

The NDCN Browser (`Omicser`) was developing using `{golem}`: "an opinionated framework for building production-grade shiny applications." - https://github.com/ThinkR-open/golem

## ETC. Caveats
Several “big” updates are planned.

+ data table visualizaiton
+ QC report tab
+ Consistent `roxygen2` headers.  Currently importing functions (e.g. `require(“package”)`) or using the `package::function` syntax is inconsistent.  
+ Migrate all dplyr::  tidyverse table manipulation to data.frame syntax and/or dtdplyr.  
(test speed improvements on mac, with caveat of apple disabled multithreading)
+ migration to the NDCN github: `ndcn/omicser`.


```{r, render-quickstart1-5, eval=TRUE, include = FALSE}
rmarkdown::render("quickstart/01_environment_setup.Rmd")
rmarkdown::render("quickstart/02_install.Rmd")
rmarkdown::render("quickstart/03_data_curation.Rmd")
rmarkdown::render("quickstart/04_configuration.Rmd")
rmarkdown::render("quickstart/05_browsing.Rmd")
```


