---
title: "1-Environment Setup"
author: "andy henrie"
date: "9/14/2021"
output:
  html_document: 
    toc: yes
    keep_md: yes
  md_document:
    variant: gfm
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Creating and running your own browser requires both Python and R.
We provide two different options for installing and setting up 
the environment you'll use when creating your browser:

- [Setup via R](#setup-via-r)
- [Setup via command line](#setup-via-command-lin)

Regardless of which approach you use,
you'll also need to install [R](https://www.r-project.org/)
and [RStudio](https://www.rstudio.com/products/rstudio/download/)
for later steps in this tutorial.
This project was developed using R version 4.1.0 and RStudio v1.4.1717.

## Setup via R

### Python

We will use a [conda environment](https://conda.io/projects/conda/en/latest/user-guide/concepts/environments.html) 
to manage Python packages handling data on the back end of the browser application.
If you don't already have **Python version 3.9** installed via Anaconda or Miniconda,
please follow [these instructions](https://conda.io/projects/conda/en/latest/user-guide/install/index.html).

## R

Once we have Python installed,
our next step will set up our R requirements,
primarily `devtools` and `reticulate`:

```{r, RM-install, eval=FALSE}
install.packages("devtools")
install.packages("reticulate")
```

> Installing the `omicser` browser package in the next tutorial will also install additional R packages as dependencies.

Now, use R to create your environment using `reticulate`:

```{r RM-r_env, eval=FALSE}
YOUR_CONDA_ENV <- "omxr"
PYTHON_VERSION <- 3.9

require(reticulate)
reticulate::install_miniconda() 
reticulate::conda_create(YOUR_CONDA_ENV, python_version = PYTHON_VERSION)
```

We'll now install Python packages into this conda environment that are required in the example scripts used later in this tutorial.
Please note that if you have bioconductor configured as a channel it may cause problems.
Forcing install from the "conda-forge" channel is preferred:

```{r RM-r_scanpy, eval=FALSE}
YOUR_CONDA_ENV <- "omxr"
reticulate::conda_install(envname= YOUR_CONDA_ENV,
                          channel = "conda-forge",
                          packages = c("scanpy", "leidenalg") )
```

If you have successfully complete all steps in this section,
you are now ready for [Installation](02_install.md). 

## Setup via command line

If you prefer to work on the command line,
the following instructions demonstrate how to use the shell to set up your environment
(examples for MacOSX):

First, download and install Miniconda:

```{bash, conda_inst, eval = FALSE}
wget https://repo.anaconda.com/miniconda/Miniconda3-latest-MacOSX-x86_64.sh -O ~/miniconda.sh
bash ~/miniconda.sh -b -p $HOME/miniconda
```

Now we can use conda to install a python 3.9 environment to use with `reticulate`. 
For this example we've called the environment `omxr`.
Installing the python packages `scanpy` and `leidenalg` will ensure all additional python requirements are installed:

```{bash, conda_env, eval = FALSE}
conda create --name omxr python=3.9 -c conda-forge scanpy leidenalg                                                                           
conda activate omxr
```

If you have successfully complete all steps in this section,
you are now ready for [Installation](02_install.md).
