---
title: "1-Installation"
date: "12/13/2021"
output:
  html_document: 
    toc: yes
    keep_md: yes
  md_document:
    variant: gfm
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Creating and running your own browser requires both Python and R.
We recommend using R to install Python and setup your environment.

> If you prefer to work on the command line, please see the troubleshooting section [below](#troubleshooting).

First ensure you have [R](https://www.r-project.org/)
and [RStudio](https://www.rstudio.com/products/rstudio/download/) installed.
This project was developed using R version 4.1.0 and RStudio v1.4.1717.

## Setting up R and Python

First, let's install the R tools we'll use to create and manage our other installations:

```{r, RM-install, eval=FALSE}
install.packages("devtools")
install.packages("reticulate")
```

> Installing the `omicser` browser package in the next section will also install additional R packages as dependencies.

We will use a [conda environment](https://conda.io/projects/conda/en/latest/user-guide/concepts/environments.html) 
to manage the Python packages handling data on the back end of the browser application. Although python virtual environments or other python installations can be configured via `reticulate` ["TODO:link to other python configurations"] we recommend using conda via the reticulate mini-conda installation as described here.

We can use R to install **Python version 3.8** and other necessary Python packages
by creating a conda environment using `reticulate`. For curation we need to make sure we also install the `scanpy` package. 

```{r RM-r_env-1, eval=FALSE}
OMICSER_PYTHON <- "my_py_env_name"

reticulate::install_miniconda() #in case it is not already installed
# simpler pip pypi install
packages <- c("scanpy[leiden]")
reticulate::conda_create(OMICSER_PYTHON, python_version = 3.8)
reticulate::conda_install(envname=OMICSER_PYTHON,
                          channel = "conda-forge",
                          pip = TRUE,
                          packages =  packages )
```

Start a fresh R-session and force reticulate to use our `OMICSER_PYTHON` by calling `reticulate::use_condaenv`.

```{r RM-r_env-2, eval=FALSE}

reticulate::use_condaenv(condaenv = OMICSER_PYTHON,
                                conda = reticulate::conda_binary(),
                                required = TRUE)

```


## Installing `omicser`

Now that we've set up our environment, we can install the package from which we'll build the browser application.
using the `devtools` package:

```{r, RM-install-1, eval=FALSE}
devtools::install_github("ndcn/omicser")
```

We are relying on this command to install not only the `omicser` package,
but a large number of additional packages that will support data formatting and app development.
If you are prompted to install and/or update any of these packages,
we recommend you do so.

> If you receive a message indicating a package failed to install (resulting in a failure to install `omicser`), please see the troubleshooting section [R packages failing to load](#r-packages-failing-to-load).

If you have successfully complete all steps in this section,
you are now ready for [data curation](02_data_curation.md). 

### _dev_ Installation

If you also intend to develop the code to contribute features to the opensource project or just to extend the functionality for you use, you should download the repo code and use the `{golem}` framework tools to help load the code locally.  
For instance _after_ cloning the repo, make sure you are in the repo directory (i.e. `omicser/`) and run:
```{r, RM-dev-install-1, eval=FALSE}
library("golem")
OMICSER_RUN_DIR <- getwd()
golem::document_and_reload(pkg = OMICSER_RUN_DIR)
```
which will load the omicser package into your R-session exposing all of the `omicser` functions.

## Troubleshooting

### command line installation

WIP.  [conda] (https://docs.conda.io/projects/conda/en/latest/user-guide/install/index.html)


### `reticulate` and `conda`

If you have difficulty loading `scanpy` (e.g., "ModuleNotFoundError: No module named 'scanpy'"),
you can check the details of your Python environment:

```{r, eval = FALSE}
reticulate::py_discover_config()
```

Because you used `reticulate` to create an environment for you 
with the relevant packages installed,
you should see your environment name
listed in the Python paths
(e.g., `r-miniconda/envs/my_py_env_name/bin/python`).

If you do not see your environment listed in the paths,
you can set the `RETICULATED_PYTHON`:

```{r, eval = FALSE}
reticulate::use_condaenv(condaenv = "my_py_env_name", required = TRUE)
```


This may also require you to restart your R session.
For more information, please see the [`reticulate` website](https://rstudio.github.io/reticulate/index.html)

### R packages failing to load

If you receive output in your console indicating one of these additional packages did not install correctly,
you can manually install it.
This may also be necessary in a later step of this series of tutorials,
if you realize a function and/or package are not available.

For example, the download for `vroom` may have stalled,
resulting in an error in your console saying the package isn't available.
You can install this package alone using the Packages interface in RStudio,
or using the following code:

```{r, RM-manual-install, eval=FALSE}
install.packages("vroom")
```

You may need to re-install and/or reload other packages as well (if they are dependent on the package you manually installed).
