---
title: "1-Installation"
date: "10/26/2021"
output:
  html_document: 
    toc: yes
    keep_md: yes
  md_document:
    variant: gfm
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Creating and running your own browser requires both Python and R.
We recommend using R to install Python and setup your environment.

> If you prefer to work on the command line, please see the troubleshooting section [Command line installation](#command-line-installation).

First ensure you have [R](https://www.r-project.org/)
and [RStudio](https://www.rstudio.com/products/rstudio/download/) installed.
This project was developed using R version 4.1.0 and RStudio v1.4.1717.

## Setting up R and Python

First, let's install the R tools we'll use to create and manage our other installations:

```{r, RM-install, eval=FALSE}
install.packages("devtools")
install.packages("reticulate")
```

> Installing the `omicser` browser package in the next section will also install additional R packages as dependencies.

We will use a [conda environment](https://conda.io/projects/conda/en/latest/user-guide/concepts/environments.html) 
to manage the Python packages handling data on the back end of the browser application.
We can use R to install **Python version 3.9** and other necessary Python packages
by creating a conda environment using `reticulate`:

```{r RM-r_env, eval=FALSE}
YOUR_CONDA_ENV <- "omxr"
PYTHON_VERSION <- 3.9

library(reticulate)
reticulate::install_miniconda() 
reticulate::conda_create(YOUR_CONDA_ENV, python_version = PYTHON_VERSION)
```

We'll now install Python packages into this conda environment that are required in the example scripts used later in this tutorial.
Please note that if you have bioconductor configured as a channel it may cause problems.
Forcing install from the "conda-forge" channel is preferred:

```{r RM-r_scanpy, eval=FALSE}
reticulate::conda_install(envname= YOUR_CONDA_ENV,
                          channel = "conda-forge",
                          packages = c("scanpy", "leidenalg") )
```

## Installing `omicser`

Now that we've set up our environment,
we can install the package from which we'll build the browser application.
We want to receive the most up-to-date version of the package possible,
so we'll install the current code from GitHub using the `devtools` package:

```{r, RM-install-1, eval=FALSE}
devtools::install_github("ndcn/omicser")
```

We are relying on this command to install not only the `omicser` package,
but a large number of additional packages that will support data formatting and app development.
If you are prompted to install and/or update any of these packages,
we recommend you do so.

> If you receive a message indicating a package failed to install (resulting in a failure to install `omicser`), please see the troubleshooting section [R packages failing to load](#r-packages-failing-to-load).

If you have successfully complete all steps in this section,
you are now ready for [data curation](02_data_curation.md). 

## Troubleshooting

### Command line installation

If you have difficulty installing Miniconda via R,
or prefer to work on the command line,
the following instructions demonstrate how to use the shell to set up your environment
(examples for MacOSX).

First, download and install Miniconda:

```{bash, conda_inst, eval = FALSE}
curl https://repo.anaconda.com/miniconda/Miniconda3-latest-MacOSX-x86_64.sh -o ~/miniconda.sh
bash ~/miniconda.sh -p $HOME/miniconda
```

You may need to respond to some prompts in your Console to complete this installation.

Now we can use conda to install a python 3.9 environment to use with `reticulate`. 
Installing the python packages `scanpy` and `leidenalg` will ensure all additional python requirements are installed:

```{bash, conda_env, eval = FALSE}
conda create -n omxr python=3.9
conda install -c conda-forge scanpy leidenalg                                                                       
```

Note that we do not use `conda activate` to load the environment.
Because we'll be working in R,
we recommend loading the environment through R
(see next section for more information).
Please also remember to install `devtools` and `reticulate`.

### `reticulate` and `conda`

If you have difficulty loading `scanpy` (e.g., "ModuleNotFoundError: No module named 'scanpy'"),
you can check the details of your Python environment:

```{r, eval = FALSE}
reticulate::py_discover_config()
```

Because you used `reticulate` to create an environment for you 
with the relevant packages installed,
you should see your environment name ("omxr")
listed in the Python paths
(e.g., `r-miniconda/envs/omxr/bin/python`).

If you do not see your environment listed in the paths,
you can load your environment:

```{r, eval = FALSE}
reticulate::use_condaenv(condaenv = "omxr", required = TRUE)
```

This may also require you to restart your R session.
For more information, please see the [`reticulate` website](https://rstudio.github.io/reticulate/index.html)

### R packages failing to load

If you receive output in your console indicating one of these additional packages did not install correctly,
you can manually install it.
This may also be necessary in a later step of this series of tutorials,
if you realize a function and/or package are not available.

For example, the download for `vroom` may have stalled,
resulting in an error in your console saying the package isn't available.
You can install this package alone using the Packages interface in RStudio,
or using the following code:

```{r, RM-manual-install, eval=FALSE}
install.packages("vroom")
```

You may need to reinstall and/or reload other packages as well (if they are dependent on the package you manually installed).
