---
title: "2-Installation"
output:
  html_document: 
    keep_md: yes
    toc: true
  md_document:
    variant: gfm
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Now that we've set up our environment,
we can install the package from which we'll build the browser application.
We want to receive the most up-to-date version of the package possible,
so we'll install the current code from GitHub using the `devtools` package:

```{r, RM-install-1, eval=FALSE}
devtools::install_github("ndcn/omicser")
```

We are relying on this command to install not only the `omicser` package,
but a large number of additional packages that will support data formatting and app development.
If you are prompted to install and/or update any of these packages,
we recommend you do so.

Now comes the crucial (and fun!) part: [curating](03_data_curation.md) your data.

## Troubleshooting

If you receive output in your console indicating one of these additional packages did not install correctly,
you can manually install it.
This may also be necessary in a later step of this series of tutorials,
if you realize a function and/or package are not available.

For example, the download for `vroom` may have stalled,
resulting in an error in your console saying the package isn't available.
You can install this package alone using the Packages interface in RStudio,
or using the following code:

```{r, RM-manual-install, eval=FALSE}
install.packages("vroom")
```

You may need to reinstall and/or reload other packages as well (if they are dependent on the package you manually installed).

