---
title: "4-Launching and browsing"
date: "12/13/2021"
output:
  html_document: 
    keep_md: yes
    toc: true
  md_document:
    variant: gfm
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Launching the browser

Once both the data and browser are appropriately formatted and configured, 
you are ready to browse your data!

Ensuring your working directory contains the `app_config.yml` file,
launch the application by executing the following line of code:

```{r, 4-browse-1, eval=FALSE}
omicser::run_defaults()
```

Your application should automatically load in a new tab/window of your default browser (tested on Chrome and Firefox).  If you prefer, there is also an exported `run_in_browser()` function which takes the database locations as arguments:

```{r, 4-browse-2, eval=FALSE}
omicser::run_in_browser(db_root = "/path/to/database/folders",
                        database_names = c(`your db name`="database_folder_name")
                        )
```

Please note that you need to keep RStudio running to work with data in your browser.

> When you are done browsing,
> you will need to close both the browser window **and** end the process running in RStudio.
> You can end the process in R by clicking on the red stop sign icon in your Console,
> or by entering `Control C` in the Console.

If you would like an overview of the features and functionality of the browser,
please view the [Browsing vignette](browsing.md).


## Troubleshooting


- **Do I have to use RStudio** No! You can run the functions from any R session. 

- **The app didn't open in my browser! What do I do?** If the app does not load automatically,
copy the URL shown into your Console (e.g., `http://127.0.0.1:6329`)
and paste it into your browser of choice. (Chrome recommended.)

- **I closed my browser window! Do I need to relaunch my app?** No, you don't need to relaunch. Follow the instructions for the question above to view your app in a browser. 
