---
title: "Browsing"
author: "andy henrie"
date: "9/14/2021"
output:
  html_document: 
    keep_md: yes
    toc: true
  md_document:
    variant: markdown_github
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# WORK IN PROGRESS:  NOT written

# _-omic_ browsing with the NDCN Browser

From your R-prompt make sure you are in the working directory with the `omicser_options.yml` spawn the app:
```{r, 4-browse-1, eval=FALSE}
omicser::run_app(options = list(launch.browser = TRUE))
```


## Browse!!

Assuming you have already loaded the omicser package, once the .yml files have been generated and teh data placed in the right directories you are good to browse!



## Overview
Congratulations you have successfully curated and spun up the data browser!

Now what can it do!  The workflow of the tool is simple:

0. WELCOME: exposition / meta-information
1. INGESTOR: choose the database 
2. SELECTOR: choose what variables and data to visualize
3. PLAYGROUND: 
  - tables
  - group and _-omic_ summary statistics
  - differential expressions
4: ETC:  qc report, dataset meta-information, publications, etc. (cellxgene)


## Example workflow

WIP:  THIS NEEDS TO BE WRITTEN.... 
