---
title: "5-Browsing"
output:
  html_document: 
    keep_md: yes
    toc: true
  md_document:
    variant: gfm
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Launching the browser

Once both the data and browser are appropriately formatted and configured, 
you are ready to browse your data!

From the Console of RStudio,
make sure your working directory contains the `omicser_options.yml` file
(you can use `setwd()` to change your working directory).
Then launch the application by executing the following line of code:

```{r, 4-browse-1, eval=FALSE}
omicser::run_app(options = list(launch.browser = TRUE))
```

This should launch your application and automatically load it in a browser window.
Please note that you need to keep RStudio running to work with data in your browser.

> When you are done browsing,
> you will need to close both the browser window **and** end the process running in RStudio.
> You can end the process in R by clicking on the red stop sign icon in your Console,
> or by entering `Control C` in the Console.

## Getting started visualizing data

The browser interface of the application is fairly intuitive.
Each of the main sections of the browser are described below,
and reference the tabs available for selection near the top of each page.

> All aspects of the browser are customizable.
> This means that there may be differences in what you see 
> compared to the defaults that are described below.
> Please consult with the person who curated your browser for more information.

### Welcome 

This page is loaded by default when launching the browser.
It includes some basic information about the browser project,
as well as a set of instructions in the lefthand side bar.

### Ingest

This page allows you to choose the database you will be browsing.
Select the database you are visualizing from the drop-down menu
(if your app is only visualizing a single dataset,
it should appear by default),
then click "Load database."

### Playground

- identify comparisons to visualize
- choose subsets of data 
- select type of visualization

### Additional info

- context for data and analysis 

### Data table

- view and filter data in tabular format
- export subset data

## An example browsing session

> **Under construction**

## Troubleshooting FAQs

- **The app didn't open in my browser! What do I do?** If the app does not load automatically,
copy the URL shown into your Console (e.g., `http://127.0.0.1:6422`)
and paste it into your browser of choice.

- **I closed my browser window! Do I need to relaunch my app?** No, you don't need to relaunch. Follow the instructions for the question above to view your app in a browser. 
